# *R* Programming Language

* <https://www.r-project.org/>\
* **programming language** & **free** software environment\
* **R**oss Ihaka and **R**obert Gentleman (1991)
* runs on major platforms (*macOS*, *Windows*, *Linux*)\
* used largely for statistical computing and graphics\
* has a large collections of packages\

## *R base* Package 
* comes with *R* installation
* contains many general functions

## *CRAN* *([C]{.ul}omprehensive [R]{.ul} [A]{.ul}rchive [N]{.ul}etwork)*
* <https://cran.r-project.org/>\
* central software repository for *R*\
* ~17,600 packages (manually reviewed)\

```{r, echo=TRUE, message = FALSE, warning=FALSE}
#install.packages("package.name")
```

## *Bioconductor*  
* <https://www.bioconductor.org/>\
* great centralized resource for NGS data analysis\
* \~2,400 packages\

```{r, echo=TRUE, message = FALSE, warning=FALSE}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#install.packages("BiocManager")
#BiocManager::install("package.name")
```

## Intorduction to Practical\

* Download the data directory from xxx\
* setwd()\
* Install and load all required *R* packages

```{r, echo=TRUE, message = FALSE, warning=FALSE}
#setwd('/Users/jordan/R_in_OMICS/input') # example
library(data.table)
```
