---
title: "R-in-OMICS"

author: Jordan Ramilowski
email: jordanr@yokohama-cu.ac.jp
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
    
date: '`r format(Sys.time(), "%d %B, %Y")`'  
---
```{r setup, echo = FALSE}
knitr::opts_knit$set(root.dir = '/Users/jordan/BI_KYOTEN/_RHomework/assignment_01/input/')
```

# *R*

* <https://www.r-project.org/>\
* **programming language** & **free** software environment\
* **R**oss Ihaka and **R**obert Gentleman (1991)
* runs on major platforms (*macOS*, *Windows*, *Linux*)\
* used largely for statistical computing and graphics\
* has a large collections of packages\

## *R base* Package 
* comes with *R* installation
* contains many general functions

## *CRAN* *([C]{.ul}omprehensive [R]{.ul} [A]{.ul}rchive [N]{.ul}etwork)*
* <https://cran.r-project.org/>\
* central software repository for *R*\
* ~17,600 packages (manually reviewed)\

```{r, echo=TRUE, message = FALSE, warning=FALSE}
#install.packages("package.name")
```

## *Bioconductor*  
* <https://www.bioconductor.org/>\
* great centralized resource for NGS data analysis\
* \~2,400 packages\

```{r, echo=TRUE, message = FALSE, warning=FALSE}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#install.packages("BiocManager")
#BiocManager::install("package.name")
```

# Intorduction to Practical\

* Download the data directory from xxx\
* setwd()\
* Install and load all required *R* packages

```{r, echo=TRUE, message = FALSE, warning=FALSE}
setwd('/Users/jordan/R_in_OMICS/input') # example
library(data.table)
```

# Reading and Writing files in *R* 

## data.table::fread\
Simple and efficient way of reading R files:\
```{r, echo=TRUE, message = FALSE, warning=FALSE}
data = fread(file = 'connectome_1in.tsv')
data = fread(file = 'connectome_1in.tsv', nThread = 4)  # can use many processors 'nThread'  
```

‘fread' vs 'read.delim'(base) speed\ 
```{r, echo=TRUE, message = FALSE, warning=FALSE}
# 'fread':
t0 = Sys.time()
data = fread(file = 'connectome_1in.tsv') 
t1 = Sys.time()-t0

# 'read.delim':
t0 = Sys.time()
data = read.delim2(file = 'connectome_1in.tsv') 
t2 = Sys.time()-t0
```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
print(c("fread: Time difference of 0.18643 secs"))
print(c("read.delim: Time difference of 4.577949 secs"))
```

## data.table::fwrite\
Simple and efficient way of writing R files:\
```{r, echo=TRUE, message = FALSE, warning=FALSE}
fwrite(data, file = 'connectome_1out.csv')             # csv file (default)
fwrite(data, file = 'connectome_1out.tsv', sep = '\t') # tab-delim file (own specification)
```

# Exploring data in R

Read in and look at first 10 columns of **connectome** data set:
```{r, echo=TRUE, message = FALSE, warning=FALSE}
connectome_1=fread(file='connectome_1in.tsv', select = 1:10) # first 10 columns
```

```{r, echo=TRUE, message = FALSE, warning=FALSE}
dim(connectome_1) # number of rows / columns
colnames(connectome_1) # names of data
```

Closer look using specific *R* packages/functions:\

## using 'str'
```{r, echo=TRUE, message = FALSE, warning=FALSE}
str(connectome_1)
```
## using 'skimr::skim'
```{r, echo=TRUE, message = FALSE, warning=FALSE}
library(skimr)
skimr::skim(connectome_1)
```

